Unshuffling Permutations

### 2015-09-27 submission.tex
typos

    (\emph{1} $\Rightarrow$ \emph{2}).
    Since $\pi$ is a square, it admits a square root $\sigma$.
    Let $2n = |\pi|$ (and hence
    $|\sigma| = n$).
    By definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
    and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
    $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
    order-isomorphic to $\sigma$.
    Let $G_\pi$ be the linear clique associated to $\pi$ with vertices
    $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
    is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
    is a subset of $\ES{G_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
    matching since $I^1 \cap I^2 = \emptyset$, and
    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$. It is also containment-free.
    Indeed, if it is not the case then there would exist two edges
    $e = \{i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$, $j < k$, in
    $\mathcal{M}$ such that $i^1_j < i^1_k < i^2_k < i^2_j$ or
    $i^1_j < i^2_k < i^1_k < i^2_j$. This is a contradiction since
    $i^2_k > i^2_j$ if $k > j$. Finally, for any two distinct edges
    $e = \{\i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
    we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
    $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing in both cases two
    elements (at position $j$ and $k$) in two patterns that are
    order-isomorphic to $\sigma$.

@@

\begin{proof}[of Proposition~\ref{proposition:hardness}]
  Suppose first that $\sigma$ occurs in $\pi$ and fix an occurrence.
  Construct a matching $\mathcal{M}$ as follows:
  \begin{itemize}
    \item $\mathcal{M}$ contains $N_1$ pairwise crossing
    $(\nu_1, \nu'_1)$-edges.
    \item $\mathcal{M}$ contains $N_2$ pairwise crossing
    $(\nu_2, \nu'_2)$-edges.
    \item $\mathcal{M}$ contains $N_3$ pairwise crossing
    $(\nu_3, \nu'_3)$-edges.
    \item $\mathcal{M}$ contains $N_4$ pairwise crossing
    $(\nu_4, \nu'_4)$-edges.
    \item $\mathcal{M}$ contains $k+2$ pairwise crossing
    $(\sigma', \pi')$-edges as depicted in
    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
    More precisely,
    (i) the first position of $\sigma'$
    (\emph{i.e.}, $(2N_1+N_2+N_3) + 1$) is linked
    to the first position of $\pi'$
    (\emph{i.e.}, $(2N_1 + 2N_2 + 2N_3 + N_4 + k + 2) + 1$),
    (ii) the last position of $\sigma'$
    (\emph{i.e.}, $(2N_1+N_2+N_3) + k+2$) is linked
    to the last position of $\pi'$
    (\emph{i.e.}, $(2N_1 + 2N_2 + 2N_3 + N_4 + k + 2) + n+2$),
    and all other positions in $\sigma'$ are linked by means of $k$ pairwise
    crossing edges to the positions in
    $\pi'$ that correspond to the fixed occurence of $\sigma$ in $\pi$.
    (Notice that we use here the fact that $\sigma$ occurs in $\pi$).
    \item $\mathcal{M}$ contains $n-k$ pairwise crossing
    $(\pi', \pi'')$-edges as depicted in
    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
    More precisely,
    all positions in $\pi'$ that do not correspond to the fixed occurence of
    $\sigma$ in $\pi$ are linked by means of $n-k$ pairwise crossing edges
    to the positions in $\pi''$ that do not correspond to the fixed
    occurrence of $\sigma$ in $\pi$.
    \item $\mathcal{M}$ contains $k$ pairwise crossing
    $(\pi'', \sigma'')$-edges as depicted in
    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
    More precisely, the positions in $\pi''$ that correspond to
    the fixed occurrence of $\sigma$ in $\pi$ are llinked
    linked by means of $k$ pairwise crossing edges to all positions in
    $\sigma''$.
    (Notice that we use here the fact that $\sigma$ occurs in $\pi$).
  \end{itemize}
  It can be easily checked (probably referring to Figure~\ref{fig:reduction}) that
  $\mathcal{M}$ is perfect, containment-free and that
  for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.

@@

  \begin{proof}[of Claim~\ref{claim:(nu_1, nu'_1)-edge}]
    Suppose, aiming at a contradiction, that there is no
    $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
    Then it follows that there exists an edge $(x, x')$ in $\mathcal{M}$,
    $1 \leq x \leq N_1$, that is neither a
    $(\nu_1, \nu_1)$-edge (since $N_1$ is odd)
    nor a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
    nor a $(\nu_1, \nu'_1)$ (by our contradiction hypothesis).
    (In other words, $x \leq N_1$ and $x' > 2N_1 + N_2$.)
    Therefore, since $\mathcal{M}$ is containment-free,
    there is neither a $(\nu_2, \nu_2)$-edge nor a $(\nu'_1, \nu'_1)$-edge
    in $\mathcal{M}$.
    Then it follows that every position $y$ of $\nu'_1$
    (\emph{i.e.}, $N_1+N_2 < y \leq 2N_1+N_2$)
    is involved in an edge $(y, y'$ of $\mathcal{M}$ with $y' \geq 2N_1+N_2$
    (otherwise $\mathcal{M}$ would not be containment-free).
    But $|\nu'_1| = N_1 = 2(2N_2 + 2n_3 + 2N_4 + 2n + 2k + 4) + 1
    > 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4$, and hence at least one
    position in $\nu'_1$ can not be involved in an edge in $\mathcal{M}$.
    Then it follows that $\mathcal{M}$ is not a perfect matching,
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

>>>>

    (\emph{1} $\Rightarrow$ \emph{2}).
    Since $\pi$ is a square, it admits a square root $\sigma$.
    Let $2n = |\pi|$ (and hence
    $|\sigma| = n$).
    By definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
    and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
    $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
    order-isomorphic to $\sigma$.
    Let $K_\pi$ be the linear clique associated to $\pi$ with vertices
    $\VS{K_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
    is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
    is a subset of $\ES{K_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
    matching since $I^1 \cap I^2 = \emptyset$, and
    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$. It is also containment-free.
    Indeed, if it is not the case then there would exist two edges
    $e = \{i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$, $j < k$, in
    $\mathcal{M}$ such that $i^1_j < i^1_k < i^2_k < i^2_j$ or
    $i^1_j < i^2_k < i^1_k < i^2_j$. This is a contradiction since
    $i^2_k > i^2_j$ if $k > j$. Finally, for any two distinct edges
    $e = \{\i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
    we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
    $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing in both cases two
    elements (at position $j$ and $k$) in two patterns that are
    order-isomorphic to $\sigma$.

@@

\begin{proof}[of Proposition~\ref{proposition:hardness}]
  Suppose first that $\sigma$ occurs in $\pi$ and fix an occurrence.
  Construct a matching $\mathcal{M}$ as follows:
  \begin{itemize}
    \item $\mathcal{M}$ contains $N_1$ pairwise crossing
    $(\nu_1, \nu'_1)$-edges.
    \item $\mathcal{M}$ contains $N_2$ pairwise crossing
    $(\nu_2, \nu'_2)$-edges.
    \item $\mathcal{M}$ contains $N_3$ pairwise crossing
    $(\nu_3, \nu'_3)$-edges.
    \item $\mathcal{M}$ contains $N_4$ pairwise crossing
    $(\nu_4, \nu'_4)$-edges.
    \item $\mathcal{M}$ contains $k+2$ pairwise crossing
    $(\sigma', \pi')$-edges as depicted in
    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
    More precisely,
    (i) the first position of $\sigma'$
    (\emph{i.e.}, $(2N_1+N_2+N_3) + 1$) is linked
    to the first position of $\pi'$
    (\emph{i.e.}, $(2N_1 + 2N_2 + 2N_3 + N_4 + k + 2) + 1$),
    (ii) the last position of $\sigma'$
    (\emph{i.e.}, $(2N_1+N_2+N_3) + k+2$) is linked
    to the last position of $\pi'$
    (\emph{i.e.}, $(2N_1 + 2N_2 + 2N_3 + N_4 + k + 2) + n+2$),
    and all other positions in $\sigma'$ are linked by means of $k$ pairwise
    crossing edges to the positions in
    $\pi'$ that correspond to the fixed occurence of $\sigma$ in $\pi$.
    (Notice that we use here the fact that $\sigma$ occurs in $\pi$).
    \item $\mathcal{M}$ contains $n-k$ pairwise crossing
    $(\pi', \pi'')$-edges as depicted in
    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
    More precisely,
    all positions in $\pi'$ that do not correspond to the fixed occurence of
    $\sigma$ in $\pi$ are linked by means of $n-k$ pairwise crossing edges
    to the positions in $\pi''$ that do not correspond to the fixed
    occurrence of $\sigma$ in $\pi$.
    \item $\mathcal{M}$ contains $k$ pairwise crossing
    $(\pi'', \sigma'')$-edges as depicted in
    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
    More precisely, the positions in $\pi''$ that correspond to
    the fixed occurrence of $\sigma$ in $\pi$ are linked
    by means of $k$ pairwise crossing edges to all positions in
    $\sigma''$.
    (Notice that we use here the fact that $\sigma$ occurs in $\pi$).
  \end{itemize}
  It can be easily checked (probably referring to Figure~\ref{fig:reduction}) that
  $\mathcal{M}$ is perfect, containment-free and that
  for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.

@@

  \begin{proof}[of Claim~\ref{claim:(nu_1, nu'_1)-edge}]
    Suppose, aiming at a contradiction, that there is no
    $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
    Then it follows that there exists an edge $(x, x')$ in $\mathcal{M}$,
    $1 \leq x \leq N_1$, that is neither a
    $(\nu_1, \nu_1)$-edge (since $N_1$ is odd)
    nor a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
    nor a $(\nu_1, \nu'_1)$ (by our contradiction hypothesis).
    (In other words, $x \leq N_1$ and $x' > 2N_1 + N_2$.)
    Therefore, since $\mathcal{M}$ is containment-free,
    there is neither a $(\nu_2, \nu_2)$-edge nor a $(\nu'_1, \nu'_1)$-edge
    in $\mathcal{M}$.
    Then it follows that every position $y$ of $\nu'_1$
    (\emph{i.e.}, $N_1+N_2 < y \leq 2N_1+N_2$)
    is involved in an edge $(y, y')$ of $\mathcal{M}$ with $y' \geq 2N_1+N_2$
    (otherwise $\mathcal{M}$ would not be containment-free).
    But $|\nu'_1| = N_1 = 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1
    > 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4$, and hence at least one
    position in $\nu'_1$ can not be involved in an edge in $\mathcal{M}$.
    Then it follows that $\mathcal{M}$ is not a perfect matching,
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}
